---
title: "Appendix"
author: ""
output: pdf_document
---

```{r setup, echo = F, message = F, warning = F}
library(tidyverse)
library(rprojroot)
library(stargazer)

knitr::opts_chunk$set(echo = F, message = F, warning = F, results='asis')

path <- find_rstudio_root_file()
```

## Online material

An interactive map of the hyperlink data and the trade flow predictions between Local Authority Districts can be found in `url removed for anonymity`.

## Out of sample $R^2$ for differnet sectors

```{r sectors}

# fig.cap="\\label{prediction}Out of sample RMSE for differnet sectors"

path.data <- paste0(path, "/data_inter/test_t2_sectors.RData")
load(path.data)

new.order <- (14:1)

rf.year.all.metrics %>%
  #round(2) %>%
  filter(metric=="Rsquared") %>%
  select(-metric) %>% 
  relocate(sector) %>%
  `colnames<-`(gsub("X", "", names(.))) %>% 
  slice(new.order) %>% 
  mutate(sector = sub('s', '', sector)) %>% 
  stargazer(summary = F,
          font.size = "footnotesize",
          type = "latex",
          #type = "text",
          header = F,
          digit.separator = ",",
          digit.separate = c(3,3),
          rownames = F,
          no.space=TRUE,
          column.sep.width = "0pt",
          notes = c("s1: Agriculture, s2: Mining, s3: Food, s4: Textiles, s5: Chemicals, s6: Equipment,", 
          "s8: Manufacturing; s9: Construction, s10: Distribution, s11: Hospitality, " , 
          "s12:	Transport, s13: Financial, s14: Real Estate, s15: Non-Market Services"),
          title = "R-squared for t + 2 out of sample predictions per sector\\label{sector_R2}"
            )
```

## Lasso regressions

```{r lasso}
rm(list=ls())
path <- find_rstudio_root_file()
path.data <- paste(path, "/data_inter/test_t2_lasso.RData", sep = "")
load(path.data)

all.metrics <- rf.year.all.metrics %>% as.data.frame() %>%
  round(2) %>%
  rownames_to_column("metric") %>%
  pivot_longer(-metric, names_to = "year", values_to="value") %>%
  pivot_wider(names_from = "metric", values_from = "value") %>%
  mutate(year = gsub("X", "", year))

# all.metrics %>% kable(format = "latex",
#       digits = 3,
#       position = "left")

stargazer(all.metrics,
          summary = F,
          font.size = "footnotesize",
          type = "latex",
          #type = "text",
          header = F,
          digit.separator = ",",
          digit.separate = c(3,3),
          rownames = F,
          no.space=TRUE,
          column.sep.width = "0pt",
#          notes = "",
          title = "LASSO regressions: accuracy metrics in unseen data from t + 2\\label{accuracy_test}")

```

## Data wrangling process

```{r examples}
path <- find_rstudio_root_file()
example1 <- read_csv(paste0(path,"/data_inter/Examples/example1.csv"))
example2 <- read_csv(paste0(path,"/data_inter/Examples/example2.csv"))
example3 <- read_csv(paste0(path,"/data_inter/Examples/example3.csv"))
example4 <- read_csv(paste0(path,"/data_inter/Examples/example4.csv"))
```

This section describes how geographic regions were added to the host-linkage dataset provided by JISC UK Web Domain Dataset^[https://data.webarchive.org.uk/opendata/ukwa.ds.2/geo/].
The archived web data are from 2000-2010.
The process begins combining the host-host links to a file containing unique postcodes for each host ending in co.uk.
An example is provided below.

```{r example1}
example1 %>% tidyr::separate(col = 3,
                      sep = "\t",
                      into = c("destination",
                               "links")) %>% 
  stargazer(summary = F,
          font.size = "footnotesize",
          type = "latex",
          #type = "text",
          header = F,
          digit.separator = ",",
          digit.separate = c(3,3),
          rownames = F,
          no.space=TRUE,
          column.sep.width = "0pt",
#          notes = "",
          title = "Host-linkage file\\label{host_link}")
```

```{r example2}
example2 %>% 
  stargazer(summary = F,
          font.size = "footnotesize",
          type = "latex",
          #type = "text",
          header = F,
          digit.separator = ",",
          digit.separate = c(3,3),
          rownames = F,
          no.space=TRUE,
          column.sep.width = "0pt",
#          notes = "",
          title = "Websites with unique postcodes\\label{unique}")
```

The two data frames were joined by matching the variable `domain`.
If an origin or destination was found in in the postcode data, it was added to the file.
Host-links without a postcode were dropped.
This leaves with host, domain and postcode for origins and destinations  and the number of links between.
This is shown below.

```{r example3}
example3 %>% dplyr::select(-X1) %>% 
  filter(!row_number() %in% 4) %>% 
  stargazer(summary = F,
          font.size = "tiny",
          type = "latex",
          #type = "text",
          header = F,
          digit.separator = ",",
          digit.separate = c(3,3),
          rownames = F,
          no.space=TRUE,
          column.sep.width = "0pt",
#          notes = "",
          title = "Combined host and postcode data\\label{combined}")
```

The next step was to remove websites that linked to themselves (e.g. the first row above).
These data were not of interest as we are looking for links between different websites.
Therefore, if origin host and destination host were the same, they were dropped.
We now have host-to-host links with the associated unique postcodes and the number of links.

The next step was to aggregate to the NUTS2 regions.
This was done by using a postcode to NUTS2 (2010 version) lookup file combined with the above created data.
The data was then aggregated summing all data with the same origin NUTS and destination NUTS codes.
We are then left with our NUTS2-to-NUTS2 links.
The same process was done for every year 2000-2010.

```{r example4}
example4 %>% stargazer(summary = F,
          font.size = "footnotesize",
          type = "latex",
          #type = "text",
          header = F,
          digit.separator = ",",
          digit.separate = c(3,3),
          rownames = F,
          no.space=TRUE,
          column.sep.width = "0pt",
#          notes = "",
          title = "NUTS2 level data\\label{nuts2}")
```

