---
title: 'Using the web to detect and predict spatial relationships'
# date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
author: |
  |
  | Emmanouil Tranos
  |
  | <small>University of Bristol, School of Geographical Sciences; Alan Turing Institute</small>
  | <small>[etranos.info](https://etranos.info/), [\@EmmanouilTranos](https://twitter.com/EmmanouilTranos), e.tranos@bristol.ac.uk</small>
output:
  revealjs::revealjs_presentation:
    theme: simple
    self_contained: true
bibliography: ../bibliography.bib
biblio-style: "apalike"
css: style.css
---

```{r, echo=FALSE, results=FALSE, include=FALSE}
library(knitr)
library(randomForest)
library(data.table)
library(stplanr)
library(dplyr)
library(ggplot2)
library(tidyr)
library(readxl)
library(httr)
library(tidyverse)
library(rgdal)
library(geosphere)
library(raster)
library(corrplot)
library(gridExtra)
library(ggrepel)
library(igraph)
library(leaflet)
library(caret)
library(DataExplorer)
library(skimr)
library(rprojroot)
#library(htmltools)
library(knitr)
library(kableExtra)
library(DescTools)
library(rprojroot)
library(doParallel)
library(tidyverse)
library(patchwork)
```

## Main idea
- Limited capacity to observe **spatial relationships**
- *Tangible*:
  - Movements of individuals
  - Vehicle or passenger flows
- A lot of work lately (transport and mobility big data)
- Various challenges regarding scale, velocity etc.

## Main idea
- *Intangible*:
  - Trade between regions
  - Business relationships
  - Knowledge flows
  - Historical ties
  - Social ties
- Difficult to capture (surveys)
- **Important** for policy designing (e.g. transport infrastructure)

## Main idea  
- Use the web to identify spatial relationships
- *Dependencies* between cities and regions
- Web text:
  - To detect and predict such relationships
  - Various spatial and temporal scales
  - *Contextualise* these relationships

# Why use the web?

## Web data and business studies
- Businesses may not expose all of their strategies on their websites, but neither do they do during surveys [@arora2013entry]
- Business websites:
    - spreading information
    - establishing a public image
    - supporting online transactions
    - sharing opinions

## Business studies using hyperlinks
<!-- Business -->
- Hyperlinks to business websites reflect business motivations and contain useful business information [@vaughan2006hyperlinks]
- Significant correlations between the number of incoming links and business performance [@vaughan2004exploring; @vaughan2004links]
- @kruger2020digital used hyperlinks between business websites in Germany to test the role of different proximity frameworks
- Innovative businesses share more hyperlinks with other business, which also tend to be innovative

## Spatial studies using hyperlinks
- Hyperlinks tend to follow national borders and gravitate towards the US [@halavais2000national]
- @kessler2017extracting used the hyperlinks between German Wikipedia webpages to represent the hierarchy of urban centres in Germany
- @salvini2016spatialization used a the English Wikipedia to build a graph of world cities
- Hyperlinks between and to administrative websites to study spatial relationships and structure [@holmberg2009local; @holmberg2010co; @janc2015geography]

## Spatial studies using hyperlinks
<!-- Blogs -->
- @lin2007blog used webblog hyperlinks to analyse the spatial reflections of the blogsphere
- @jones2010blog focused on the New York City theatre scene to investigate the existence and role of a 'virtual buzz'

# Our current work

---

- JISC UK Web Domain Dataset: .uk webpages in the Internet Archive 1996-2010
- @tranos2020individual
- @tranos2020digital
- @tranos2020hyperlinks

---

```{r , echo=FALSE, results='hide', message=FALSE}
#, fig.height=15, fig.width=10}

# This is the project path
path <- find_rstudio_root_file()

# load test_t2.RData, which was created by test_ts.Rmd
path.rdata <- paste0(path, "/data_inter/test_t2.RData")
load(path.rdata)

# create a list of the above model outputs
fit.model.all <- mget(ls(pattern = "^model.all20"))

# remove model.all2011
fit.model.all$model.all2011 <- NULL

# resamples
fit.model.all.res <- resamples(fit.model.all)

# plot RMSE

# scales <- list(x=list(relation="free"), y=list(relation="free"))
# bwplot(fit.model.all.res, scales=scales)

fit.model.all.res$values %>% #extract the values
  dplyr::select(1, ends_with("RMSE")) %>% #select the first column and all columns with a name ending with "RMSE"
  gather(model, RMSE, -1) %>% #convert to long table
  mutate(model = sub("~RMSE", "", model)) %>% #leave just the model names
  mutate(model = gsub("model.all", "", model)) %>%
  #all.metrics$year <- gsub("X", "", all.metrics$year) # couldn't put the gsub in the chain
  ggplot()+ #call ggplot
  geom_boxplot(aes(x = RMSE, y = model)) -> RMSE #and plot the box plot

RMSE <- RMSE +
  ylab("")

# RMSE <- RMSE +
#   scale_y_discrete(limits = c("model.all2001", "model.all2002", "model.all2003",
#                               "model.all2004", "model.all2005", "model.all2006",
#                               "model.all2007", "model.all2008", "model.all2009",
#                               "model.all2010"))

# plot MAE
# fit.model.all.res$values %>% #extract the values
#   dplyr::select(1, ends_with("MAE")) %>% #select the first column and all columns with a name ending with "RMSE"
#   gather(model, MAE, -1) %>% #convert to long table
#   mutate(model = sub("~MAE", "", model)) %>% #leave just the model names
#   mutate(model = gsub("model.all", "", model)) %>%
#   ggplot()+ #call ggplot
#   geom_boxplot(aes(x = MAE, y = model)) -> MAE #and plot the box plot
#
# MAE <- MAE +
#   ylab("")
# # MAE <- MAE +
# #   scale_y_discrete(limits = c("model.all2001", "model.all2002", "model.all2003",
# #                               "model.all2004", "model.all2005", "model.all2006",
# #                               "model.all2007", "model.all2008", "model.all2009",
# #                               "model.all2010"))
#
# # plot Rsquared
# fit.model.all.res$values %>% #extract the values
#   dplyr::select(1, ends_with("Rsquared")) %>% #select the first column and all columns with a name ending with "RMSE"
#   gather(model, Rsquared, -1) %>% #convert to long table
#   mutate(model = sub("~Rsquared", "", model)) %>% #leave just the model names
#   mutate(model = gsub("model.all", "", model)) %>%
#   ggplot()+ #call ggplot
#   geom_boxplot(aes(x = Rsquared, y = model)) -> Rsquared #and plot the box plot
#
# Rsquared <- Rsquared +
#   ylab("Year")
# # Rsquared <- Rsquared +
# #   scale_y_discrete(limits = c("model.all2001", "model.all2002", "model.all2003",
# #                               "model.all2004", "model.all2005", "model.all2006",
# #                               "model.all2007", "model.all2008", "model.all2009",
# #                               "model.all2010"))
#
# grid.arrange(
#   Rsquared, MAE, RMSE,
#   nrow = 1, ncol = 3,
#   top = "Accuracy metrics")
```

```{r echo=FALSE, results= 'hide', message=FALSE}

all.metrics <- rf.year.all.metrics %>% as.data.frame() %>%
  round(2) %>%
  rownames_to_column("metric") %>%
  pivot_longer(-metric, names_to = "year", values_to="value") %>%
  pivot_wider(names_from = "metric", values_from = "value") %>%
  mutate(year = gsub("X", "", year))

all.metrics %>% kable(format = "html",
      digits = 3,
      position = "left")

```

```{r echo=FALSE, results= 'markup', message=FALSE}
# there are 2 columns named year, so I drop one
# pred.by.year.all$year <- NULL

ggplot(data = pred.by.year.all,aes(x = io, y = predictions)) +
  geom_point(colour = "blue") +
  geom_abline(intercept = 0, slope = 1, colour = "red") +
  #geom_vline(xintercept = 23, colour = "green", linetype = "dashed") +
  facet_wrap(~ year,ncol = 2) +
  #coord_cartesian(xlim = c(0,50000),ylim = c(0,50000)) +
  ggtitle("Predicted vs. observed interregional trade by year") +
  ylab("Predicted") + xlab("Observed")

```

# Gap
- Importance of understanding spatial relationship for urban and regional policy making
- Some work on extracting spatial relations from the web
- Most of it is based hyperlinks between geolocated websites
- Limited work based on toponym co-occurrences [@meijers2019using]
- **Very limited** work on contextualisation of such relationships
- **Hardly any work** in utilising such data to create meaningful geographic knowledge
- Importance: inform city and regional **policies**

# The new PhD project
- ML, NLP statistical and network modelling
- Named entity recognition (actors, organisations, events, toponym disambiguation)
- Extract spatial relationships
- Contextualise (NLP)
- Multiscalar approach
- Spatio-temporal networks
- Need for methods incorporating space & text
- Tools to access large web archives

# The new PhD project
- Untapped potential to utilise the web as a data source
- Data science tools to support social science and policy making

```{css echo=FALSE, message=FALSE}
.slide {
    height: 750px;
    overflow-y: auto !important;
}

# it aligns all the slides
# .reveal section {
#         text-align: left;
#     }

```

# References {.allowframebreaks}
